Wireless service accounting based on Ethereum state channels - Master Thesis

### 2019-10-12 Paper.tex
Add backend description, disclaimer, license and correct some references

@@

In addition, depending on the different levels of
access permission, block chains can
split into two types:
1) public (such as Bitcoin and Ethereum); Y
2) private (such as Hyperledger). Blockchain serves
as a platform for smart contracts
(hereinafter, \ textit {smart contracts}).
For example, programmed in the Solidity language of
Ethereum, stay and run.
Blockchain is a technology concept
DLT distributed ledger
(\ textit {distributed ledger technology}).
It can be integrated into multiple business areas.

@@

To keep users' freedom of interrupt the service consumption
whenever they want and also avoid the overhead of transacting
many times with the smart contract authors have implemented
what is known as a state channel in a smart contract; this
makes possible keep control of the time fraction of the service
that's already consumed and at the same time cryptographically
guarantee that the smart contract is going to unlock the funds
when the provider desires to do so. State channels are a
standard pattern in most blockchain applications and in all cases
(\textit{independently of the complexity}) present three main
steps: opening, transaction and closing; project's implementation
described below:

@@

\begin{itemize}
\item[] 1) The user associates the laptop with the WiFi hotspot.
\item[] 2) Once connected is redirected to the captive portal front-end.
\item[] 3) Front provisions MetaMask as Web3 provider.
\item[] 4) Front-end uses MetaMaskto allow the user
  to send value to the contract
\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
  on port 8545.
\item[] 7) Once transaction is confirmed, the back-end iptables controller
  start forwarding user's traffic.
\end{itemize}

\subsubsection{Channel transactions}
Before sending the initial value transaction (\textit{in proportion with
  the duration of the service}) user's front-end generates an ephemeral elliptic curve
secp256k1 keypair and sends the public key to the contract to open the state channel.
Corresponding ephemeral private key is the one that's going to be used
to periodically (each 10 s approximately)
produce increasing value signatures proportional to the
service consumption ellapsed time. This signatures are being send to the
backend instead of the blockchain; backend verifies and validates the
signatures and stores the last one (\textit{the most valuable}).

@@

\paragraph{ChannelData} Includes payer ethereum address,
its state channel ephemeral address, deposit amount, opening time
and a boolean representing the open/close status of the channel.

@@

\paragraph{channelCount} uint256 variable that increments
in a unit every time a user opens a channel. This operation
does not rely on SafeMath since it's virtually impossible to
overflow a uint256 storage record.

@@

\paragraph{channelMapping} Holding a key-value map between
channelCount values and channelData structures.

@@

\paragraph{pricePerSecond} Constant keeping track of
service price per second.

@@

\paragraph{openChannel} receives as parameters, the ephemeral address
of the user and also the purchased service value.

@@

\paragraph{closeChannel}

@@

As an example, in figure 5 it's shown how three virtual operators over
different infrastructures shared or not in a slice-model of a neutral host
infrastructure; a slice can be described as a dynamic virtualization
allowing multitenancy of some physical resources like access points,
physical switches and datacenters including mobile edge computing
appliances; this is the case of currently in execution projects like
5GCity\cite{5gcity} where a distributed cloud
and radio platform for municipalities and infrastructure owners acting
as 5G neutral hosts is being developed and tested. Platforms
targeting 5G business models such as 5GCity can benefit from adding
this smart contract business logic in order to bill end-users and automate
all the accounting part.

@@

To take into account the problem illustrated in previous paragraph
we can take as a basis a model of self-sovereign identity\cite{allen}
and in particular the one recently anounced by the authorities
of Catalonia government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that the credential can be linked
to some administration database that could be queried only by a judge
requirement and guarantee the users identities are recognized

@@

\begin{itemize}
\item Fully automated accounting and billing layer.
\item Regulation compliance without KyC on-boarding process.
\item GDP compliance by default. Operator does not need to know
  user's name, bank account, etc.
\end{itemize}

@@

  This model can be extended without limits to the full wireless
  telecommunication market tokenization; as depicted in figure 6,
  from the spectrum auctioning, including infrastructure providers
  virtualizing their appliances for virtual operators, end to users
  service provision and accounting relying on a self-sovereign
  identity system.

@@

\bibitem{state-channel-contract-a}
  \href{state-channel-contract-a on GitHub}{https://github.com/ethereum-internet-access/state-channel-contract-a}
\bibitem{etherwall}
  \href{etherwall on GitHub}{https://github.com/ethereum-internet-access/etherwall}
\bibitem{etherdoor}
  \href{etherdoor on GitHub}{https://github.com/ethereum-internet-access/etherdoor}

\bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}
\bibitem{allen}  \href{Christopher Allen, The path to self sovereign identity, 2016}
  {http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
\bibitem{identicat}
  \href{Catalonia to Build DLT-Based Identity, Coindesk, 2019}{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}

\bibitem{Coinbase19} Coinbase. Intercambio de moneda digital \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}
\bibitem{ConsenSys19} Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.
\bibitem{Docker19} Docker Documentation. How nodes work. 6 Septiembre 2019.  \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}
\bibitem{ETHBerlinZwei19} Hackathon ETH Berlín Zwei. Twitter. 6 Septiembre 2019. \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}
\bibitem{Hegedus18} P. Hegedus. "Towards Analyzing the Complexity Landscape of Solidity Based Ethereum Smart Contracts". 2018 IEEE/ACM 1st International Workshop on Emerging Trends in Software Engineering for Blockchain (WETSEB). 27 Mayo-3 Junio 2018. \href{https://ieeexplore.ieee.org/document/8445056}{https://ieeexplore.ieee.org/document/8445056}
\bibitem{MakerDAO19} MakerDAO. Stability for the blockchain.  \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}
\bibitem{Munoz18} J. L. Muñoz. "Docker Containers". Apuntes de las clases magistrales. Máster en Tecnologías Blockchain. 1, Edición. UPC School. 29 Noviembre 2018.
\bibitem{Nakamoto08} S. Nakamoto. "Bitcoin: A Peer-to-Peer Electronic Cash System". 31 Octubre 2008.  \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}
\bibitem{RaspberryPi19} Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.
\bibitem{RaspberryPiBuy19} Raspberry Pi FAQs. BUYING AND SHIPPING. Where can I buy a Raspberry Pi?. Página Web oficial. 11 Agosto 2019.  \href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{https://www.raspberrypi.org/help/faqs/\#buyingWhere}
\bibitem{Stallings04} W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. Páginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
\end{thebibliography}
\clearpage

>>>>

\vspace*{5cm}

\begin{center}
\Huge{\textbf{DISCLAMER}}
\vspace{1cm}

\Large
All software described and linked in this document is open source
under the MIT license; so anyone can freely use it and make
modifications to the source code to fulfill their use
cases. But if you do so, please take into account that the software
at the time of this writing is merely
a proof of concept and is not properly audited; although
authors tried to follow best practices regarding smart contract,
front-end and back-end design and implementation there are still
some tradeoffs and recognizable security issues
described in this document's
security considerations section. Please take this into account
when using and/or modifying this software and be conscious
that authors do not provide any guarantee in case of
security issues, economical loses or regulation infringement.
\end{center}

\vspace{1cm}

\vspace*{5cm}

\begin{verbatim}
MIT License

Copyright (c) 2019 ethereum-internet-access

Permission is hereby granted, free of charge, to any person obtaining a copy
of this software and associated documentation files (the "Software"), to deal
in the Software without restriction, including without limitation the rights
to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
copies of the Software, and to permit persons to whom the Software is
furnished to do so, subject to the following conditions:

The above copyright notice and this permission notice shall be included in all
copies or substantial portions of the Software.

THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
SOFTWARE.
\end{verbatim}

@@

In addition, depending on the different levels of
access permission, block chains can
split into two types:
1) public (such as Bitcoin and Ethereum); and
2) private (such as Hyperledger). Blockchain serves
as a platform for smart contracts
(hereinafter, \textit{smart contracts}).
For example, programmed in the Solidity language of
Ethereum, stay and run.
Blockchain is a technology concept
DLT distributed ledger
(\textit{distributed ledger technology}).
It can be integrated into multiple business areas.

@@

To keep users' freedom to interrupt the service consumption
whenever they want and also avoid the overhead of transacting
many times with the smart contract; authors have implemented
what is known as a state channel in a smart contract; this
makes possible to keep control of the time fraction of the service
that's already consumed and at the same time cryptographically
guarantee that the smart contract is going to unlock the funds
when the provider desires to do so. State channels are a
standard pattern in most blockchain applications and in all cases
(\textit{independently of the complexity}) present three main
steps: opening, transaction and closing; project's particular implementation
described below:

@@

\begin{itemize}
\item[] 1) The user associates the laptop with the WiFi hotspot.
\item[] 2) Once connected is redirected to the captive portal front-end.
\item[] 3) Front provisions MetaMask as Web3 provider.
\item[] 4) Front-end uses MetaMask to allow the user
  to send value to the contract
\item[] 5-6) Signed transaction is relayed through a light Infura proxy listening
  on port 8545.
\item[] 7) Once transaction is confirmed, the back-end iptables controller
  start forwarding user's traffic.
\end{itemize}

\subsubsection{Channel transactions}
Before sending the initial value transaction (\textit{proportional to
  the initially expected duration of the service})
user's front-end generates an ephemeral elliptic curve
secp256k1 keypair and sends the public key to the contract to open the state channel.
Corresponding ephemeral private key is the one that's going to be used
to periodically (each 10 s approximately)
produce increasing value signatures proportional to the
service consumption ellapsed time. This signatures are being send to the
backend instead of the blockchain; backend verifies and validates the
signatures and stores the last one (\textit{the most valuable}).

@@

\paragraph{\texttt{ChannelData}} Includes payer ethereum address,
its state channel ephemeral address, deposit amount, opening time
and a boolean representing the open/close status of the channel.

@@

\paragraph{\texttt{channelCount}} uint256 variable that increments
in a unit every time a user opens a channel. This operation
does not rely on SafeMath since it's virtually impossible to
overflow a uint256 storage record.

@@

\paragraph{\texttt{channelMapping}} Holding a key-value map between
channelCount values and channelData structures.

@@

\paragraph{\texttt{pricePerSecond}} Constant keeping track of
service price per second.

@@

\paragraph{\texttt{openChannel}} receives as parameters, the ephemeral address
of the user and also the purchased service value.

@@

\paragraph{\texttt{closeChannel}} accepts the amount of ETH to refund
to sender, the \texttt{channelId} and the \texttt{signature} of the
payer; checks the signature and after updating the state, transfers
the amount to the owner of the contract and the return to the channel
opener.

\paragraph{\texttt{claimTimeout}} method that allows the payer to
close the channel in case contract owner does not close it after
some \texttt{expirationTime}.

\subsubsection{Etherwall}

Consists on a NodeJS backend that fulfils three main system functions:

\begin{itemize}
\item Controls the state of the current connections.
\item iptables interfacing.
\item Micro-payment signature and amount validations.
\end{itemize}

On execution uses the module \texttt{lib/iptables.js}
to setup and initialize system's iptables redirecting
all 80 port traffic to the ip address to Raspberry's
ip address; to do so, relyes on the node module \texttt{child\_process}
to spawn system commands. Also redirects all traffic trying
to reach 8545 port to itself the same way in order to
redirect MetaMask's RCP calls to the lightweight infuraProxy\cite{infuraProxy}
developed also for this project and contained in etherwall's root directory.

After initializing, the backend using the web framework\cite{express}
etherwall exposes the following endpoints:

\paragraph{\texttt{GET /generate\_204}} As a way to enable
captive portal detection to Android OS. HTTP Status 204
means the file exists but, is empty.
This lets Android know that the Internet is accessible.
If the request receives HTTP Status 302 (temporary redirect)
instead of HTTP 204, Android will follow the
redirection URL to display the Captive Portal to the user.

\paragraph{\texttt{POST /mac}} Endpoint to request a new
connection; mainly
accepts the time left to use the service as a parameter
and internally gets the MAC address of user's device
making an arp table query based on client's ip address.
After that invokes \texttt{lib/iptables.js} \texttt{grantAccess}
function in order to insert a rule to enable packet forwarding
for this particular MAC address on top of iptables.
Also uses module \texttt{lib/connections.js} to keep persistence
of the connection status. Once connected after posting
\texttt{timeLeft} to this endpoint, in module
\texttt{lib/connections.js} there's a \texttt{setInterval}
triggering getting connection status and time since last
micro-payment each 5 seconds. In case \texttt{timeSinceLastPayment}
exceeds 60 seconds or \texttt{timeLeft} goes negative,
disconnects the client deleting the forwarding rule from
iptables.

\paragraph{\texttt{GET /mac}} Retrieves the status of client's
connection from \texttt{lib/connections.js} module; this module
in addition, provides some serialized persistence using a JSON
file under the \texttt{data/} directory to keep state integrity
between server restarts.

\paragraph{\texttt{POST /payment}} Endpoint that allows the front-end
to send a payload containing a payload passing the corresponding
contract data:

\begin{itemize}
\item \texttt{channelId}
\item \texttt{amount} (to unlock)
\item \texttt{signature}
\end{itemize}

In order to verify the correctness of the payload, uses Web3
module to compute the \texttt{soliditySha3} and compares it
to the signature's \texttt{messageHash}; after that recovers
the public address from the signature and compares it to
the associated smart contract mapping structure based on its
\texttt{channelId}; if everything matches, it additionally checks
that the unlocked amount is approximately what it corresponds
taking into account connection ellapsed time
(\textit{using a kind threshold of 60 seconds marginal cost});
in case everything is correct, stores the payment together
with a timestamp as the last one corresponding
to the connection using again \texttt{lib/connections.js} module.
This is done to allow bootstrapped \texttt{setInterval}
on the previously described \texttt{POST mac/} to process
the timeout business logic.

\subsubsection{etherdoor}

@@

As an example, in figure 5 it's shown how three virtual operators over
different infrastructures shared or not in a slice-model of a neutral host
infrastructure; a slice can be described as a dynamic virtualization
allowing multitenancy of some physical resources like access points,
physical switches and datacenters including mobile edge computing
appliances; this is the case of currently ongoing projects like
5GCity\cite{5gcity} where a distributed cloud
and radio platform for municipalities and infrastructure owners acting
as 5G neutral hosts is being developed and tested. Platforms
targeting 5G business models such as 5GCity can benefit from adding
this smart contract business logic in order to bill end-users and automate
all the accounting layer.

@@

To take into account the problem illustrated in previous paragraph
we can take as a basis a model of self-sovereign identity\cite{allen}
and in particular the one recently anounced by the authorities
of Catalonian government\cite{identicat} that will provide credentials
to citizens using DLT technologies as a Public Key Infrastructure; this
model could allow in some near future to give the citizen the option
of registering identity in such a way that the credential can be linked
to some administration database that could be queried only by a judge
requirement and guarantee the users identities are recognized

@@

\begin{itemize}
\item Fully automated accounting and billing layer.
\item Regulation compliance without KyC on-boarding process.
\item GDPR compliance by default. Operator does not need to know
  user's name, bank account, etc.
\end{itemize}

@@

  This model can be extended without limits towards a full wireless
  telecommunication market tokenization; as depicted in figure 6,
  from the spectrum auctioning, including infrastructure providers
  virtualizing their appliances for virtual operators, end to users
  service provision and accounting relying on a self-sovereign
  identity system.

@@

\bibitem{state-channel-contract-a}
  \href{https://github.com/ethereum-internet-access/state-channel-contract-a}{state-channel-contract-a on GitHub}
\bibitem{etherwall}
  \href{https://github.com/ethereum-internet-access/etherwall}{etherwall on GitHub}
\bibitem{infuraProxy}
  \href{https://github.com/ethereum-internet-access/etherwall/blob/master/infuraProxy.js}{infuraProxy on GitHub}

\bibitem{etherdoor}
  \href{https://github.com/ethereum-internet-access/etherdoor}{etherdoor on GitHub}
\bibitem{express}
  \href{https://expressjs.com/}{https://expressjs.com/}

\bibitem{5gcity} \href{https://www.5gcity.eu/}{https://www.5gcity.eu/}
\bibitem{allen}  \href{http://www.lifewithalacrity.com/2016/04/the-path-to-self-soverereign-identity.html}
  {C. Allen, The path to self sovereign identity, 2016}
\bibitem{identicat}
  \href{https://www.coindesk.com/catalonia-government-to-build-dlt-based-identity-platform-for-citizens}
       {Catalonia to Build DLT Identity, Coindesk, 2019}

\bibitem{Coinbase19} Coinbase. Intercambio de moneda digital \href{https://www.coinbase.com/price/ethereum}{https://www.coinbase.com/price/ethereum}
\bibitem{ConsenSys19} Ethereum Smart Contract Best Practices. Security tools. ConsenSys. 12 Agosto 2019.
\bibitem{Docker19} Docker Documentation. How nodes work. 6 Septiembre 2019.  \href{https://docs.docker.com/engine/swarm/how-swarm-mode-works/nodes/}{Docker Swarm}
\bibitem{ETHBerlinZwei19} Hackathon ETH Berlín Zwei. Twitter. 6 Septiembre 2019. \href{https://twitter.com/ETHBerlin}{https://twitter.com/ETHBerlin}
\bibitem{Hegedus18} P. Hegedus. "Towards Analyzing the Complexity Landscape of Solidity Based Ethereum Smart Contracts". 2018 IEEE/ACM 1st International Workshop on Emerging Trends in Software Engineering for Blockchain (WETSEB). 27 Mayo-3 Junio 2018. \href{https://ieeexplore.ieee.org/document/8445056}{https://ieeexplore.ieee.org/document/8445056}
\bibitem{MakerDAO19} MakerDAO. Stability for the blockchain.  \href{https://makerdao.com/en/dai}{https://makerdao.com/en/dai}
\bibitem{Munoz18} J. L. Muñoz. "Docker Containers". Apuntes de las clases magistrales. Máster en Tecnologías Blockchain. 1, Edición. UPC School. 29 Noviembre 2018.
\bibitem{Nakamoto08} S. Nakamoto. "Bitcoin: A Peer-to-Peer Electronic Cash System". 31 Octubre 2008.  \href{https://bitcoin.org/bitcoin.pdf}{https://bitcoin.org/bitcoin.pdf}
\bibitem{RaspberryPi19} Raspberry Pi 3 Modelo B - Placa Base (1,2 GHz Quad-Core Arm Cortex-A53, 1 GB RAM, USB 2.0). Amazon. 11 Agosto 2019.
\bibitem{RaspberryPiBuy19} Raspberry Pi FAQs. BUYING AND SHIPPING. Where can I buy a Raspberry Pi?. Página Web oficial. 11 Agosto 2019.  \href{https://www.raspberrypi.org/help/faqs/#buyingWhere}{https://www.raspberrypi.org/help/faqs/\#buyingWhere}
\bibitem{Stallings04} W. Stallings. "Comunicaciones y redes de computadores". 1 Septiembre 2004. Páginas 904. Pearson Prentice Hall. ISBN-13: 978-8420541105.
\end{thebibliography}
\clearpage
